Discrete Quantum Theories and Computing

### 2019-04-21 dissertation.tex
dissertation.lyx "Chapter 6 Conclusion and Further Discussion" Updating...

Although it is convenient to assume physical quantities are as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick is made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
be an integer multiple of the elementary charge. A perfect square
cannot be drawn physically. Even we don't draw it, but merely ask
whether it exists in the space. This question is still beyond our
ability to answer because whether space itself is ultimately discrete
or continuous is still an unsolved question. Even if a perfect square
exists physically, we may not able to decide whether the ``square''
we discussed is really perfect or almost perfect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision in reality is limited generally to three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Hence, when the quantities used to branch the computation closes to
zero, the branch chosen by a BCSS machine cannot reliably predict
the branch chosen by a realistic analog computer. Since two branches
of computations may not have any relations, the follow-up computational
paths may be significantly different, and this difference is unlikely
to be compensated by error analysis techniques easily like the butterfly
effect. Therefore, we might not be able to utilize a realistic analog
computer to acquire the computational power predicted by this kind
of theoretical models.%
\begin{comment}
\todo{Maybe extended to consider probabilistic model? }
\end{comment}

@@

Conventional quantum theory is based on the continuum of complex numbers,
but we cannot distinguish two arbitrary complex numbers without unbounded
resources. To explore alternative versions of quantum theory incorporating
our limitation of distinguishability, two types of discrete quantum
theories were described: \emph{quantum theories and computing over
finite fields} and \emph{quantum interval-valued probability measures
(QIVPMs)}. Examining the computational and physical consequences of
such frameworks could yield new insights into not only the subtle
properties of conventional quantum theory but also the power and capacity
of quantum computing.

The theories over finite fields started with unrestricted discrete
fields (Chapter~\ref{chap:Unrestricted-Finite-Fields}), then advanced
to a more reasonable framework based on complexifiable discrete fields
(Sec.~\ref{discretequantumtheoryI} to~\ref{discretequantumcomputingI}),
which lacked a notion of probability and supported unnaturally efficient
deterministic quantum algorithms. A still more plausible discrete
theory with cardinal probabilities was proposed (Sec.~\ref{sec:Discrete-Quantum-Theory-(II)}
and~\ref{discretequantumcomputingII}), where conventional quantum
theory and computing emerged in a local sense, but lacking arithmetic
operations among cardinal probabilities still posed difficulty to
define expectation values. Since the axiomatic approach looked unlikely
to provide ``sensible'' real-valued probability measures over finite
fields (Sec.~\ref{sec:Toward-IVPM}), we shifted our attention to
directly embed our limitation of distinguishability into the theory
to define QIVPM.

\begin{figure}[b]
\centering{}$\xymatrix{\textrm{Classical Probability Measure}\ar[d]_{\textrm{blur probability}}\ar[r]^{\textrm{glue events}} & \textrm{Quantum Probability Measure}\ar[d]^{\textrm{blur probability}}\\
\textrm{Classical IVPM}\ar[r]_{\textrm{glue events}} & \textrm{QIVPM}
}
$\caption{\label{fig:commutative-diagrams}QIVPMs inherited from both quantum
probability measures and classical IVPMs.\todo{The first arrow is
too short!}}
\end{figure}
As a natural extension of both convention quantum probability measures
and classical interval-valued probability measures (IVPMs) illustrated
in Fig.~\ref{fig:commutative-diagrams}, QIVPMs inherited definitions
and properties from the both sides. While the expectation values of
QIVPMs could be pulled back to the classical cases and consistent
with those of quantum probability measures in the infinitely precise
limit (Sec.~\ref{sec:Interval-Uncertainty}), foundational concepts
in quantum mechanics, such as the Kochen-Specker and Gleason theorems,
extended to QIVPMs in subtle ways. By carefully specifying experimental
uncertainties, we established rigorous bounds on the validity of the
Kochen-Specker theorem (Sec.~\ref{sec:Kochen-Specker}). While there
is a QIVPM not consistent with Gleason's unique state~$\rho$ on
all projectors, we constructed a class of QIVPMs for which the original
Gleason theorem could be recovered asymptotically (Sec.~\ref{sec:Gleason}).

In the following further discussion, we will briefly explain why we
can only recover Gleason's on a class of QIVPM, the possibility to
further build a computational model over QIVPMs, and the possibility
to combine two approaches to consider quantum interval-valued probability
measures over finite fields.

@@

$\frameR\colon\CP{2}\rightarrow\left[0,1\right]$ defined by $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]=\bar{\mu}\left(\proj{\psi}\right)$

@@

A successful interval-valued theory might be further extended to the
quantum theory over finite fields based on the following definition.

>>>>

Although it is convenient to assume physical quantities are as continuous
as their mathematical models, each physical quantity discussed in
the previous paragraph could not have a simple one-to-one correspondence
with a real number. Chuang-tzu's stick is made by molecules and cannot
be cut in halves endlessly. The input charge of an integrator must
be an integer multiple of the elementary charge. A perfect square
cannot be drawn physically. Even we don't draw it, but merely ask
whether it exists in the space. This question is still beyond our
ability to answer because whether space itself is ultimately discrete
or continuous is still an unsolved question. Even if a perfect square
exists physically, we may not able to decide whether the ``square''
we discussed is really perfect or almost perfect differing only by
an arbitrarily small extent. Similar situation applying to the length
of a slide rule, it is beyond our ability to decide whether its length
is ultimately discrete or continuous. Even if its length is continuous,
identifying which real number represents its length requires infinite
precision, but the precision in reality is limited generally to three
or four significant figures~\cite{wiki:AnalogComputer}. As the results,
even if analog computers really store real numbers, they cannot be
precisely read, written, and used to branch the computation. The opposite
of the last statement is sometimes assumed by real computational models.
For example, Blum, Cucker, Shub, and Smale modeled their BCSS machine
to allow the operations of deciding whether a number is greater or
equal to zero over real numbers and whether a number is exactly zero
over complex numbers \cite{BSS1989,Ziegler2007,blum2012complexity}.
Hence, when the quantities used to branch the computation closes to
zero, the branch chosen by a BCSS machine cannot reliably predict
the branch chosen by a realistic analog computer. Since two branches
of computations may not have any relations, the follow-up computational
paths may be significantly different, and this difference is unlikely
to be compensated by error analysis techniques easily like the butterfly
effect. Therefore, we might not be able to utilize a realistic analog
computer to acquire the computational power predicted by this kind
of theoretical models.%
\begin{comment}
\todo{Maybe extended to consider probabilistic model? For BCSS machine,
maybe using the idea of BPP to define the complexity class is more
reasonable than P or NP?}
\end{comment}

@@

Conventional quantum theory is based on the continuum of complex numbers,
but we cannot distinguish two arbitrary complex numbers without unbounded
resources. To explore alternative versions of quantum theory incorporating
our limitation of distinguishability, two types of discrete quantum
theories were described: \emph{quantum theories and computing over
finite fields} and \emph{quantum interval-valued probability measures
(QIVPMs)}. Examining the computational and physical consequences of
such frameworks can yield new insights into not only the subtle properties
of conventional quantum theory but also the power and capacity of
quantum computing.

The theories over finite fields started with unrestricted discrete
fields (Chapter~\ref{chap:Unrestricted-Finite-Fields}), then advanced
to a more reasonable framework based on complexifiable discrete fields
(Secs.~\ref{discretequantumtheoryI} to~\ref{discretequantumcomputingI}),
which lacks a notion of probability and supports unnaturally efficient
deterministic quantum algorithms. A still more plausible discrete
theory with cardinal probabilities was proposed (Secs.~\ref{sec:Discrete-Quantum-Theory-(II)}
and~\ref{discretequantumcomputingII}), where conventional quantum
theory and computing emerge in a local sense, but lacking arithmetic
operations among cardinal probabilities still posed difficulty to
define expectation values. Since the axiomatic approach looked unlikely
to provide sensible real-valued probability measures over finite fields
(Sec.~\ref{sec:Toward-IVPM}), we shifted our attention to directly
embed our limitation of distinguishability into the theory to define
QIVPM.

\begin{figure}[b]
\centering{}$\xymatrix{\textrm{Classical Probability Measure}\ar[d]_{\textrm{blur probability}}\ar[rr]^{\textrm{glue events}} &  & \textrm{Quantum Probability Measure}\ar[d]^{\textrm{blur probability}}\\
\textrm{Classical IVPM}\ar[rr]_{\textrm{glue events}} &  & \textrm{QIVPM}
}
$\caption{\label{fig:commutative-diagrams}QIVPMs inherit from both quantum
probability measures and classical IVPMs.}
\end{figure}
As a natural extension of both conventional quantum probability measures
and classical interval-valued probability measures (IVPMs) illustrated
in Fig.~\ref{fig:commutative-diagrams}, QIVPMs inherit definitions
and properties from the both sides. While the expectation values of
QIVPMs can be pulled back to the classical cases and consistent with
those of quantum probability measures in the infinitely precise limit
(Sec.~\ref{sec:Interval-Uncertainty}), foundational concepts in
quantum mechanics, such as the Kochen-Specker and Gleason theorems,
extended to QIVPMs in subtle ways. By carefully specifying experimental
uncertainties, we established rigorous bounds on the validity of the
Kochen-Specker theorem (Sec.~\ref{sec:Kochen-Specker}). While there
is a QIVPM not consistent with Gleason's unique state~$\rho$ on
all projectors, we constructed a class of QIVPMs for which the original
Gleason theorem could be recovered asymptotically (Sec.~\ref{sec:Gleason}).

In the following further discussion, we will briefly explain why we
only recovered Gleason's on a class of QIVPMs, the possibility to
further build a computational model over QIVPMs, and the possibility
combining both approaches to consider quantum interval-valued probability
measures over finite fields.

@@

$\frameR\colon\CP{2}\rightarrow\left[0,1\right]$ defined by $[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)]=\bar{\mu}\left(\proj{\psi}\right)$

@@

A successful interval-valued theory might be further extended over
finite fields based on the following definition.
